Apache Flink\texttrademark: Unified Stream and Batch Processing \\ in a Single Engine

### 2015-11-26 streaming.tex
incorporate changes in stateful stream processing and windowing

\subsection{Stateful Stream Processing}
\TODO{Asterios: This is very developer-heavy. We need to show the big picture. Long story short, the side-effect free functional view of the system changes with mutable state that is managed by the system. Why do we need state? What can be state? How is it implemented in Flink. @Paris, please answer these 3 questions in 2-3 paragraphs. The example below could also be missing.}

While most operators in Flink’s DataStream API look like functional, side-effect free operators, they also support efficient stageful computations. Explicit mutable state is incorporated in the Flink DataStream API by providing: $i)$ Operator interfaces or annotations to statically register explicit local variables within the scope of an operation;  $ii)$ an OperatorState abstraction for marking partitioned key-value states and the associated operations. Users can also configure how the state is stored and checkpointed using the StateBackend abstractions provided by the system allowing very flexible custom state management in the streaming applications. Flink’s checkpointing mechanism guarantees that any registered state is durably accessed with exactly-once update semantics.

@@

\subsection{Stream Windows}
Many incremental computations over unbounded streams are often evaluated over finite logical views, called windows. Apache Flink creates windows via a highly flexible process which is composed out of three core functions: a window assigner and optionally a trigger and an evictor. All three functions can be selected among a pool of common predefined implementations (e.g. sliding time window) or explicitly defined by the user (via UDFs).

>>>>

\subsection{Stateful Stream Processing}
While most operators in Flink’s DataStream API look like functional, side-effect free operators, they also support efficient stateful computations. State is critical to many applications such as machine learning model building, graph analysis, user session handling and window aggregations among others. There is a plethora of different types of states depending on the use case. For example, the state can be something as simple as a counter or a sum, or more complex, such as a classification tree or a large sparse matrix often used in machine learning applications. Stream windows can also be modelled, in principle, by stateful operators which assign records to continuously updated buckets kept in memory as part of an operator state.

In Apache Flink, state is made explicit, to cover all different use cases, and incorporated in the API by providing: $i)$ Operator interfaces or annotations to statically register explicit local variables within the scope of an operation;  $ii)$ an OperatorState abstraction for marking partitioned key-value states and the associated operations. Users can also configure how the state is stored and checkpointed using the StateBackend abstractions provided by the system allowing very flexible custom state management in the streaming applications. Flink’s checkpointing mechanism guarantees that any registered state is durably accessed with exactly-once update semantics.

@@

\subsection{Stream Windows}
Many incremental computations over unbounded streams are often evaluated over finite logical views, called windows. Apache Flink incorporates windowing within a stateful operator which is configured via a flexible declaration composed out of three core functions: a window assigner and optionally a trigger and an evictor. All three functions can be selected among a pool of common predefined implementations (e.g. sliding time window) or explicitly defined by the user (via UDFs).
